The \frb\ Campaign: Multi-Observatory Radio Burst and Implications for the FRB Population

### 2017-03-23 repeater-vladata.tex
refining frb rate discussion

The calculation of a volumetric rate has been used to identify classes of transient and their defining physical processes \citep{2006ARA&A..44..507W, 2012Sci...337..927G}. \citet{2013Sci...341...53T} used this approach to estimate an isotropic rate of $10^{-3}$\ galaxy$^{-1}$\ yr$^{-1}$ by assuming a characteristic redshift horizon of 0.9. This rate is high \citep[comparable to the rate of core-collapse supernovae;][]{2006Natur.439...45D}, which has been used to argue that FRBs are likely not associated with other classes of transient, such as GRBs or rarer SNe.

This conclusion is in tension with new models inspired by multi-wavelength associations to \frb. Its coincidence with a compact, persistent radio source is consistent with \frb as a young magnetar in a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb is found in a low-metallicity dwarf galaxy, which is the preferred environment for long GRBs and Type-I superluminous supernovae \citep[LGRB and SLSN-I, respectively;][]{2017arXiv170102370M}.

We now reevaluate this rate using the properties of \frb\ as a prototype FRB. First, we recast this estimate as a beamed, volumetric rate, $R_{FRB} = R_p /(\Omega_b V (z))$, where $R_p$\ is the projected FRB rate, $\Omega_b$\ is the beaming fraction \citep[$\sim0.1$;][]{1998MNRAS.298..625T} and $V(z)$\ is the volume out to redshift $z$. Second, the projected FRB rate is now believed to be lower than initially believed \citep[$2\times10^3\ \rm{sky}^{-1} \rm{day}^{-1}$\ at high Galactic latitudes and flux densities brighter than 1 Jy ms; ][]{2016arXiv161100458L, 2016MNRAS.460L..30C, 2016MNRAS.455.2207R}. Including these effects, we estimate a volumetric rate of $6\times10^{-5} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$, assuming a typical FRB horizon redshift of 0.9. For comparison, the LGRB and SLSN-I rates are orders of magnitude lower, near $10^{-8}$ Mpc$^{-3}$\ yr$^{-1}$ \citep{2012Sci...337..927G}.

The revised FRB rate is an improvement, but it still ignores multiple effects. The burst spectral structure discussed here shows that this estimate likely underestimates the FRB rate by missing bursts that fall outside the observing band (typically a few hundred MHz near 1.4~GHz). For a typical burst envelope width of 500~MHz centered at GHz-scale frequencies, this could underestimate the rate by a factor of a few. The horizon scale ($z=0.9$) is not well defined and can change the implied rate by an order of magnitude.

sensitivity is the biggest awkward factor!
an out: use luminosity distribution for absolute rate scale?

calculate number of repetitions and compare to Metzger limit

%DM dominated by the IGM \citep[$DM\approx z\times1200 \rm{pc}\ \rm{cm}^{-3}$ for fully ionized hydrogen or $DM\approx z\times900 \rm{pc}\ \rm{cm}^{-3}$ for more realistic ionization state;][]{2003ApJ...598L..79I,2004MNRAS.348..999I}. 
%galaxy number density $\Phi(M)$
%They also calculated the number of galaxies by assuming a characteristic $L_*$\ galaxy \citep[corresponding to stellar mass $M_* \approx 1011 M$;][]{2012MNRAS.421..621B}.
%
%This isotropic FRB rate assumes a projected FRB rate at a single (observationally defined) flux limit and that no bursts repeat. Using this calculation with data from more sensitive telescopes, we might infer a higher rate. However, if we assume that \frb\ is representative of the overall FRB population, then more sensitive observations would be more likely to find repeating FRBs that would effectively depress the estimate of the underlying isotropic rate per source. In this model, these two effects partially cancel, although without a physical model it is difficult to say how well.

>>>>

The calculation of a volumetric rate has been used to identify classes of transient and their defining physical processes \citep{2006ARA&A..44..507W}. \citet{2013Sci...341...53T} used this approach to estimate an isotropic rate of $10^{-3}$\ galaxy$^{-1}$\ yr$^{-1}$ by assuming a characteristic redshift horizon of 0.9 and a typical Milky Way-like galaxy host. This rate is high \citep[comparable to the rate of core-collapse supernovae;][]{2006Natur.439...45D}, which has been used to argue that FRBs are likely not associated with other classes of transient, such as GRBs or rarer SNe.

This conclusion is in tension with new models inspired by multi-wavelength associations to \frb. Its coincidence with a compact, persistent radio source is consistent with \frb as a young magnetar in a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb is found in a low-metallicity dwarf galaxy, which is the preferred environment for long GRBs and Type-I superluminous supernovae \citep[LGRB and SLSN-I, respectively;][]{2017arXiv170102370M}.
**biased tracer of star formation**

We now reevaluate this rate using the properties of \frb\ as a prototype FRB. First, we recast this estimate as volumetric rate of FRB sources by including beaming and repetition effects. The volumetric source rate is $R_{FRB} = R_p /(N_r \Omega_b V (z))$, where $R_p$\ is the projected FRB rate, $N_r$\ is the number of bursts per source, $\Omega_b$\ is the beaming fraction \citep[$\sim0.1$;][]{1998MNRAS.298..625T} and $V(z)$\ is the volume out to redshift $z$. Second, the projected FRB rate is now believed to be lower than initially believed \citep[$2\times10^3\ \rm{sky}^{-1} \rm{day}^{-1}$\ at high Galactic latitudes and flux densities brighter than 1 Jy ms; ][]{2016arXiv161100458L, 2016MNRAS.460L..30C, 2016MNRAS.455.2207R}.

% alternatively
% Third, the measured energy distributions for \frb\ allow us to estimate their detectability out to the distances implied by their DMs.
% 
% what is the energy distribution of known FRBs assuming their DM measures distance?

Third, the largest published DM \citep[$\sim1500$\ pc cm$^{-3}$ and much higher values rumored;][]{2016MNRAS.460L..30C} may imply a larger cosmic volume for FRBs. Assuming a DM dominated by the IGM \citep[$DM\approx z\times900 \rm{pc}\ \rm{cm}^{-3}$ for realistic ionization state;][]{2003ApJ...598L..79I,2004MNRAS.348..999I}, this places a maximum FRB redshift on the order of 2, near the peak the in the cosmic star formation history \citep{2014ARA&A..52..415M}. While this is consistent with magnetar models proposed for \frb, the DM may include significant contributions from the host galaxy and intervening galaxy halos \citep{2014ApJ...780L..33M}. Using an assumed horizon redshift of 2, we estimate a rate of $R_p = 10^{-5} N_r^{-1} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$.

This revised FRB rate still ignores multiple effects. Narrow burst spectral structure implies that observations underestimate by missing bursts that fall outside the observing band. For a typical burst envelope width of 500~MHz centered at GHz-scale frequencies, past observations (typically using bandwidths of a few hundred MHz near 1.4~GHz) would underestimate the rate by a factor of a few. The horizon scale is not well defined and can change the implied rate by a factor of $\sim5$.

For comparison, the LGRB and SLSN-I rates are orders of magnitude lower, near $10^{-8}$ Mpc$^{-3}$\ yr$^{-1}$ \citep{2012Sci...337..927G}.

%galaxy number density $\Phi(M)$
%They also calculated the number of galaxies by assuming a characteristic $L_*$\ galaxy \citep[corresponding to stellar mass $M_* \approx 1011 M$;][]{2012MNRAS.421..621B}.
%
%This isotropic FRB rate assumes a projected FRB rate at a single (observationally defined) flux limit and that no bursts repeat. Using this calculation with data from more sensitive telescopes, we might infer a higher rate. However, if we assume that \frb\ is representative of the overall FRB population, then more sensitive observations would be more likely to find repeating FRBs that would effectively depress the estimate of the underlying isotropic rate per source. In this model, these two effects partially cancel, although without a physical model it is difficult to say how well.
